---
title: "SDS 4030 - Statistics for Data Science II"
subtitle: Final Project Code
author: 
output: pdf_document
fontsize: 12pt
header-includes:
  - \usepackage{hyperref}
  - \hypersetup{pdfpagemode={UseNone}, pdfstartview=}
  - \usepackage{amsthm}
  - \theoremstyle{definition}
  - \newtheorem{problem}{Problem}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(fig.align = "center", fig.height = 4, fig.width = 5)
knitr::opts_chunk$set(tidy.opts = list(width.cutoff = 68), tidy = TRUE)
knitr::opts_chunk$set(message = FALSE)
set.seed(0)
```

\problem
\textit{Proof}.
```{r}
# loading dataset Here

horse= read.table("horse.txt", header = TRUE)
head(horse)

```

```{r}
#fitted model w only quant variables

quant_model = lm(respiratory_rate ~ temperature + pulse + packed_cell_volume + total_protein, data = horse)
summary(quant_model)
```

```{r}
# qual, exact, large sample size -- lecture 21 reference


beta <- coef(quant_model)
X <- model.matrix(quant_model)
Y = horse$respiratory_rate
n <- nrow(X)
p <- ncol(X)
S2 <- sum(residuals(quant_model)^2) / (n - p) 


R <- matrix(c(
  0,  5,   0,    0, 350, 
  0,  0,  40,  100,   0 
), nrow = 2, byrow = TRUE)

r <- c(3, 3)
k <- length(r)

FT <- crossprod(R %*% beta - r, solve(R %*% solve(t(X) %*% X) %*% t(R), R %*% beta - r)) / (k * S2)
print(FT)

pvalue <- pf(FT, k, n - p, lower.tail = FALSE)
print(pvalue)

quant_reduced <- lm(respiratory_rate ~ I(temperature + total_protein + pulse + packed_cell_volume), data=horse)
summary(quant_reduced)


anova(quant_reduced, quant_model)


```

```{r}
# qual, asymptotic, large sample size -- lecture 21 reference

WT = k * FT
print(WT)
pvalue = pchisq(WT, k, lower.tail = FALSE)
print(pvalue)
anova(quant_reduced, quant_model, test = "Chisq")
```

The conclusions agree of both the exact and asymptotic tests agree with each other. Both had a p-value of 0.9043546, so we fail to reject the null hypothesis. The Wald statistic and Exact F statistic are related as 

\[W_n = \frac{\sigma}{S^2_{nQ1}} = k F\]

where F is the F statistic from the exact test. So we would expect the conclusions of both tests to agree with each other.


\qed

\problem
\textit{Proof}.

```{r}
#jackknife
betajack <- matrix(0, n, p)
for (i in 1:n) {
  jack <- lm(respiratory_rate ~ temperature + pulse + packed_cell_volume + total_protein,data = horse[-i, ])
  betajack[i, ] <- coef(jack)
}
betabar <- colMeans(betajack)
print(betabar)

SE <- apply(betajack, 2, sd) * sqrt((n - 1)^2 / n)
print(SE)

```

```{r}
#nonparametric bootstrap
nboot <- 10000
betaboot <- matrix(0, nboot, p)
set.seed(123)

for (k in 1:nboot) {
  indices <- sample(1:n, n, replace = TRUE)
  boot <- lm(respiratory_rate ~ temperature + pulse + packed_cell_volume + total_protein, data = horse[indices, ])
  betaboot[k, ] <- coef(boot)
}

betabar <- colMeans(betaboot)
print(betabar)

SE <- apply(betaboot, 2, sd)
print(SE)

```

```{r}
alpha <- 0.05

CInormal <- cbind(betabar - qnorm(1 - alpha / 2) * SE,
                  betabar + qnorm(1 - alpha / 2) * SE)
print(CInormal)

CIpercentile <- cbind(
  apply(betaboot, 2, quantile, probs = alpha / 2),
  apply(betaboot, 2, quantile, probs = 1 - alpha / 2)
)
print(CIpercentile)

beta_full <- coef(quant_model)
CIpivotal <- cbind(
  2 * beta_full - apply(betaboot, 2, quantile, probs = 1 - alpha / 2),
  2 * beta_full - apply(betaboot, 2, quantile, probs = alpha / 2)
)
print(CIpivotal)

```
```{r}
summary(quant_model)$coefficients[, "Std. Error"]

```
The coefficient estimates are essentially the same across methods (OLS / jackknife mean / bootstrap mean). Both resampling methods give very different estimates than OLS. The Jackknife SEs and bootstrap SEs are very similar. These SEs are much larger than OLS SEs. The two sets of intervals are very similar to one another, further supporting the stability of the bootstrap results. Compared to OLS confidence intervals, the bootstrap intervals are significantly wider.

\qed

\problem
\textit{Proof}.

```{r}
# Nonparametric test of staistical significant: Permutation F Test

f = summary(quant_model)$fstatistic[1]
print(f)
pvalue = pf(f, p, n - p - 1, lower.tail = FALSE)
print(pvalue)
nperm = 10000
fperm = numeric(nperm)
for (k in 1:nperm)
{
  Yperm = sample(Y)
  perm = lm(respiratory_rate ~ temperature + pulse + packed_cell_volume + total_protein, horse)
  fperm[k] = summary(perm)$fstatistic[1]
}
pvalue = (1 + sum(fperm > f)) / (1 + nperm)
print(pvalue)
```
The estimated F test p-value through the permutation method is equal to 1e-4. Since the pvalue is less than alpha, we can reject the null hypothesis. The pvalue under the theoretical null distribution is 6.4e-12, which is smaller than the pvalue from the permutation test. However, they both lead to the rejection of the null.

```{r}
# Nonparametric test of statistical significance for the effect of temperature
# Permutation t Test

t = summary(quant_model)$coefficients["temperature", 3]
print(t)
pvalue = 2 * pt(abs(t), n - p - 1, lower.tail = FALSE)
print(pvalue)
aux = lm(respiratory_rate ~ pulse + packed_cell_volume + total_protein, horse)
fitted = aux$fitted.values
residuals = aux$residuals
tperm = numeric(nperm)
for (k in 1:nperm)
{
Yperm = fitted + sample(residuals)
perm = lm(Yperm ~ temperature + pulse + packed_cell_volume + total_protein, horse)
tperm[k] = summary(perm)$coefficients["temperature", 3]
}
pvalue = (1 + sum(abs(tperm) > abs(t))) / (1 + nperm)
print(pvalue)

```
The estimated t test p-value through the permutation method is equal to 0.017. Since the pvalue is less than alpha, we can reject the null hypothesis. The pvalue under the theoretical null distribution is 0.015, which is very similar to the pvaqlue obtained from the permutation test.

\qed

\problem
\textit{Proof}.

```{r}
# Check for possible covariate transformations

par(mfrow = c(1, 5))
boxplot(horse$temperature, ylab = "Temperature", pch = 16)
boxplot(sqrt(horse$temperature), ylab = "Square Root of Temperature", pch = 16)
boxplot(log(horse$temperature), ylab = "Log of Temperature", pch = 16)
boxplot((horse$temperature)^2, ylab = "Square of Temperature", pch = 16)
boxplot((horse$temperature)^(-1), ylab = "Inverse of Temperature", pch = 16)

boxplot(horse$pulse, ylab = "Pulse", pch = 16)
boxplot(sqrt(horse$pulse), ylab = "Square Root of Pulse", pch = 16)
boxplot(log(horse$pulse), ylab = "Log of Pulse", pch = 16)
boxplot((horse$pulse)^2, ylab = "Square of Pulse", pch = 16)
boxplot((horse$pulse)^(-1), ylab = "Inverse of Pulse", pch = 16)

boxplot(horse$total_protein, ylab = "Total Protein", pch = 16)
boxplot(sqrt(horse$total_protein), ylab = "Square Root of Total Protein", pch = 16)
boxplot(log(horse$total_protein), ylab = "Log of Total Protein", pch = 16)
boxplot((horse$total_protein)^2, ylab = "Square of Total Protein", pch = 16)
boxplot((horse$total_protein)^(-1), ylab = "Inverse of Total Protein", pch = 16)

boxplot(horse$packed_cell_volume, ylab = "Packed Cell Volume", pch = 16)
boxplot(sqrt(horse$packed_cell_volume), ylab = "Square Root of Packed Cell Volume", pch = 16)
boxplot(log(horse$packed_cell_volume), ylab = "Log of Packed Cell Volume", pch = 16)
boxplot((horse$packed_cell_volume)^2, ylab = "Square of Packed Cell Volume", pch = 16)
boxplot((horse$packed_cell_volume)^(-1), ylab = "Inverse of Packed Cell Volume", pch = 16)
```
Pulse and total protein seem skewed, so need transformations.
For pulse, the best transformation is inverse. For total protein, the best transformation is a log transformation. A log transformation is reasonable for packed cell volume to decrease outliers and make for a more symmetric distribution.

```{r}
fit = lm(respiratory_rate ~ temperature + (1/pulse) + log(packed_cell_volume) + log(total_protein), horse)
plot(fit$fitted.values, rstandard(fit), xlab = "Fitted Values",
ylab = "Standardized Residuals", pch = 16)
abline(h = 0, lty = 2, lwd = 2)
```
Since constant variance is violated, a transformation should be applied to the response variable.

```{r} 
#Box-Cox transformation
horse$inv_pulse = 1 / horse$pulse
Y = horse$respiratory_rate

lambda = seq(-2, 2, 0.01)
loglik = numeric(401)
for (i in 1:401)
{
  if (lambda[i] == 0)
  {
    Ypower = log(Y)
  }
  else
  {
    Ypower = (Y^lambda[i] - 1) / lambda[i]
  }
  power = lm(Ypower ~ temperature + inv_pulse + log(packed_cell_volume) + log(total_protein), horse)
  loglik[i] = logLik(power)[1] + (lambda[i] - 1) * sum(log(Y))
}

CI = range(lambda[loglik > max(loglik) - qchisq(alpha, 1, lower.tail = FALSE) / 2])
print(CI)
```
Since the CI is close to covering 0, a log transformation is reasonable.

```{r}

power = lm(log(respiratory_rate) ~ temperature + inv_pulse + log(packed_cell_volume) + log(total_protein), horse)
plot(power$fitted.values, rstandard(power), main = "Log-
Transformation", xlab = "Fitted Values", ylab = "
Standardized Residuals", pch = 16)
abline(h = 0, lty = 2, lwd = 2)

library(qqconf)
qq_conf_plot(rstandard(power), points_params = list(pch = 16, cex = 0.5))

```
All standardized residuals fall within the band, except one. This is enough to reject normality.

\qed

\problem
\textit{Proof}.
```{r}

#Best subset selection with valid post-selection inference 

library(leaps)

#New horse data frame with 5 transformed variables
horse_transformed = data.frame(
  Y = log(horse$respiratory_rate),
  temperature = horse$temperature,
  inv_pulse = 1 / horse$pulse,
  log_pcv  = log(horse$packed_cell_volume),
  log_tp   = log(horse$total_protein)
)

Y = horse_transformed$Y
X =  horse_transformed[, -1]      
n = nrow(X)
P = ncol(X)


#Set.seed for consistency
set.seed(111)

train = sample(n, 0.6 * n)
valid = sample(setdiff(1:n, train), 0.2 * n)
test  =  setdiff(1:n, c(train, valid))
MSPE = numeric(P)

full = lm(Y ~ ., horse_transformed, train)
sub = regsubsets(formula(full), horse_transformed[train, ], nvmax = P)
summary(sub)$which

for (p in 1:P)
{
  fit = lm(Y ~ ., X[, summary(sub)$which[p, -1], drop = FALSE], train)
  preds = predict(fit, X[valid, summary(sub)$which[p, -1], drop = FALSE])
  MSPE[p] = mean((Y[valid] - preds)^2)
}

plot(MSPE, type="b", pch=16, lwd=4, ylab="Validation MSPE")

p =  which.min(MSPE)
final_model = lm(Y ~ ., X[, summary(sub)$which[p, -1], drop = FALSE], test)
summary(final_model)


```

Based on the validation MSPE curve, the full (4 covariates) model provides the most accurate predictor for respiratory rate. Among all of the predictors, inverse pulse (inv_pulse) is the only statistically significant variable on the test set and shows that horses with higher pulse rates tend to have higher respiratory rates. The remaining predictors (temperature, log(packed cell volume), and log(total protein)) are not individually statistically significant.

\qed

\problem
\textit{Proof}.
```{r}
#Investigate collinearity 
library(faraway)
cor(horse_transformed)
plot(horse_transformed, pch = 16, cex = 0.25)

cat("VIF:", vif(full))

```

Based on the VIFs and correlation plots, the transformed quantitative variables aren't very collinear. 
```{r}
#Ridge Regression 
library(glmnet)
train_boot = c(train, valid)
X = as.matrix(X)
cv.ridge = cv.glmnet(X[train_boot,], Y[train_boot], nfolds = 6, alpha = 0)
ridge = glmnet(X[test,], Y[test], alpha = 0, lambda = cv.ridge$lambda.min)
print(ridge$beta)

#Bootstrap
nboot = 10000
beta = matrix(0, nboot, P)
colnames(beta) = colnames(X)
for (i in 1:nboot)
{
  ind = sample(test, replace = TRUE)
  beta[i,] = glmnet(X[ind,], Y[ind], alpha = 0, lambda = cv.ridge$lambda.min)$beta[,]
}
SE = apply(beta, 2, sd)
cat("SE after Bootstrap:", SE, "\n")


ols = lm(Y ~ ., horse_transformed[train, ])
cat("Summary of OLS coefficients for comparison:","\n")
summary(ols)$coefficients


```

The ridge regression coefficients all shrunk closer to zero in comparison to the OLS coefficients. The bootstrap results SEs are slightly higher than the SEs calculated using OLS. 

\qed

\problem
\textit{Proof}.

```{r}
#Lasso Regression
cv.lasso = cv.glmnet(X[train,], Y[train], alpha = 1)
lasso = glmnet(X[test,], Y[test], nfolds = 6, alpha = 1, lambda
= cv.lasso$lambda.min)
print(lasso$beta)

for (i in 1:nboot)
{
  ind = sample(test, replace = TRUE)
  beta[i,] = glmnet(X[ind,], Y[ind], alpha = 1, lambda = cv.lasso$lambda.min)$beta[,]
}
SE = apply(beta, 2, sd)
cat("SE after Bootstrap:", SE, "\n")
```
While some of the estimated lasso coefficients become extremely close to zero, others stay the same. Because the quantitative variables aren't collinear, the estimated SEs of the lasso coefficients are greater than the SEs estimated by the OLS. 
\qed

\problem
\textit{Proof}.

```{r}
#Investigate the interaction effects between pain and all available quantitative covariates on the response variable.

Y = log(horse$respiratory_rate)
inv_pulse = 1 / horse$pulse
log_pcv  = log(horse$packed_cell_volume)
log_tp   = log(horse$total_protein)
horse$pain = as.factor(horse$pain)


fit_interact = lm( Y ~ pain*(temperature + inv_pulse + log_pcv + log_tp), horse)
summary(fit_interact)

anova(fit_interact)

```
```{r}
fit_interact = update(fit_interact, . ~ . - pain:log_tp)
summary(fit_interact)
anova(fit_interact)
```

```{r}
fit_interact = update(fit_interact, . ~ . - pain:log_pcv)
summary(fit_interact)
anova(fit_interact)
```

```{r}
fit_interact = update(fit_interact, . ~ . - log_pcv)
summary(fit_interact)
anova(fit_interact)
```

```{r}
fit_interact = update(fit_interact, . ~ . - log_tp)
summary(fit_interact)
anova(fit_interact)
```

Final Model Equation: 
\[
\begin{aligned}
Y_i
&= \beta_0
+ \beta_1\,\text{pain}_{i,\text{continuous}}
+ \beta_2\,\text{pain}_{i,\text{depressed}}
+ \beta_3\,\text{pain}_{i,\text{mild}}
+ \beta_4\,\text{pain}_{i,\text{severe}} \\
&\quad + \beta_5\,\text{temperature}_i
+ \beta_6\,\text{inv\_pulse}_i \\
&\quad + \gamma_1\,\text{pain}_{i,\text{continuous}}\cdot\text{temperature}_i
+ \gamma_2\,\text{pain}_{i,\text{depressed}}\cdot\text{temperature}_i
+ \gamma_3\,\text{pain}_{i,\text{mild}}\cdot\text{temperature}_i
+ \gamma_4\,\text{pain}_{i,\text{severe}}\cdot\text{temperature}_i \\
&\quad + \delta_1\,\text{pain}_{i,\text{continuous}}\cdot\text{inv\_pulse}_i
+ \delta_2\,\text{pain}_{i,\text{depressed}}\cdot\text{inv\_pulse}_i
+ \delta_3\,\text{pain}_{i,\text{mild}}\cdot\text{inv\_pulse}_i
+ \delta_4\,\text{pain}_{i,\text{severe}}\cdot\text{inv\_pulse}_i
+ \varepsilon_i.
\end{aligned}
\]


Interpretation of each covariate in the final model: 

- paincontinuous = –12.1040
Holding other predictors constant, a horse classified as continuous pain is expected to have a log respiratory rate about 12.10 units lower than a horse with no pain.

- paindepressed = –17.5964
Holding other predictors constant, a depressed horse has a log respiratory rate about 17.60 units lower than a horse with no pain.

- painmild = –10.8063
Holding other predictors constant, a horse with mild pain has a log respiratory rate about 10.81 units lower than a no-pain horse.

- painsevere = –13.0211
Holding other predictors constant, a horse with severe pain has a log respiratory rate about 13.02 units lower than a no-pain horse.

-temperature = –0.2557
For a horse in the no-pain category with a fixed pulse, a 1°C increase in temperature is associated with a 0.256-unit decrease in log respiratory rate. Temperature is significant overall (ANOVA p = 0.0227), but its slope for no-pain horses is not individually significant.

- inv_pulse = –64.7889
Because inv_pulse = 1/pulse, a decrease in inv_pulse corresponds to a higher pulse rate. For a horse in the no-pain category, holding temperature constant, a one-unit increase in inv_pulse is associated with a 64.79-unit decrease in log respiratory rate.

- paindepressed:temperature = 0.4473
For depressed horses, the effect of temperature on log respiratory rate is 0.447 units more positive per degree than for no-pain horses.This is the only significant interaction in the model (p = 0.0151).

\qed

\problem
\textit{Proof}.
```{r}
#Mixed-effects Model

Y = log(horse$respiratory_rate)
horse$pain = factor(horse$pain)
horse$peristalsis = factor(horse$peristalsis)
horse$abdominal_distension = factor(horse$abdominal_distension)
horse$peripheral_pulse = factor(horse$peripheral_pulse)
horse$temperature_of_extremities = factor(horse$temperature_of_extremities)
horse$surgery <- factor(horse$surgery)
horse$outcome <- factor(horse$outcome)
```

```{r}
#chosen random effects

# Used transformed variables from problem 8 for reference

library(lme4)
mixed_eff_mod = lmer(
  Y ~ temperature + inv_pulse + log_pcv + log_tp +
(1 | temperature_of_extremities) +(1 | abdominal_distension) +(1 | surgery),
  data = horse)

```

```{r}
sigma = data.frame(summary(mixed_eff_mod)$varcor)$sdcor
names(sigma) = data.frame(summary(mixed_eff_mod)$varcor)$grp
```

```{r}
ICC = sigma^2 / sum(sigma^2)
print(ICC)
```

```{r}
summary(mixed_eff_mod, correlation = FALSE)
anova(mixed_eff_mod)
```
The covariates chosen as random effects are abdominal_distension, surgery and temperature_of_extremities, since these were the categorical variables with the largest nonzero variance components in the mixed-effects model. All categorical variables with exactly zero variance components were treated as fixed effects, as instructed.

As we determined in question 4, the log transformation is the suitable transformation for our response variable.

After fitting a mixed-effects model and examining the variance components, three categorical covariates had nonzero variances: abdominal_distension, temperature_of_extremities, and surgery.

The total variation in your response variable explained by each random effect is as follows: abdominal_distension (variance = 0.02668, 2.67%) and temperature_of_extremities (variance = 0.00652, 0.65%), and surgery (variance = 0.005579, 0.56%).


Another interpretation for each ICC:
ICC(abdominal_distension) = 0.0936
Two horses with the same abdominal distension level have an expected correlation of about 9.4% in their log respiratory rates. This points too a moderate grouping effect of adominal distension level on respiratory rate.

ICC(temperature_of_extremities) = 0.0229
Two horses with the same extremity temperature category have an expected correlation of about 2.3% in their log respiratory rates. This points to a very small grouping effect of extremity temperature on respiratory rate.

ICC(surgery) = 0.01956768
Two horses that underwent the same type of surgery have an expected correlation of about 2.0% in their log respiratory rates. This points to a very small grouping effect of surgery on respiratory rate.



\qed

\problem
\textit{Proof}.
```{r}
base_model <- lm(Y ~ temperature + inv_pulse + log_pcv + log_tp, data = horse)
print(summary(model))
```

```{r}
Y = log(horse$respiratory_rate)
X = model.matrix(base_model)
n = dim(X)[1]
p = dim(X)[2] - 1
residuals = base_model$residuals
plot(rstandard(base_model) ~ temperature + inv_pulse + log_pcv + log_tp, horse, ylab = "Standardized
Residuals", pch = 16)
abline(h = 0, lty = 2, lwd = 2)
```
```{r}
#Studentized Breusch-Pagan Test
Yaux = residuals(base_model)^2
aux = lm(Yaux ~ temperature + inv_pulse + log_pcv + log_tp, horse)
BP = n * summary(aux)$r.squared
print(BP)
pvalue = pchisq(BP, p, lower.tail = FALSE)
print(pvalue)
bptest(base_model)
```

```{r}
#Brown-Forsythe Test -- pain

library(car)
leveneTest(Y ~ pain, data = horse)
```

```{r}
#Brown-Forsythe Test -- outcome
library(car)
leveneTest(Y ~ outcome, data = horse)

```
```{r}
#Brown-Forsythe Test -- temp of extremities

library(car)
leveneTest(Y ~ temperature_of_extremities, data = horse)

```

```{r}
#Brown-Forsythe Test -- peripheral pulse
library(car)
leveneTest(Y ~ peripheral_pulse, data = horse)
```

```{r}
#Brown-Forsythe Test -- abdominal distension

library(car)
leveneTest(Y ~ abdominal_distension, data = horse)

```

```{r}
#Brown-Forsythe Test -- peristalsis

library(car)
leveneTest(Y ~ peristalsis, data = horse)
```
For the quantitative variables, a studentized Breusch–Pagan test was conducted. Because the p-value is well above 0.05, we fail to reject the null hypothesis of constant variance. So here is no evidence of heteroscedasticity, and the residual variance does not appear to depend on any of the included quantitative predictors.


For the qualitative varibles, a Brown–Forsythe tests was conducted. 

The Brown–Forsythe test yields p = 0.0006161, providing statistically significant evidence that the residual variance differs across pain groups (heteroscedasticity).

The Brown–Forsythe test yields p = 0.2814, so there is no statistically significant evidence of differing variances across outcome groups (homoscedasticity).

The Brown–Forsythe test yields p = 0.4136, so there is no statistically significant evidence that the residual variance differs across peripheral-pulse categories.

The Brown–Forsythe test yields p = 0.0444, indicating statistically significant evidence that residual variances differ across abdominal-distension groups (heteroscedasticity).

The Brown–Forsythe test yields p = 0.061, so there is no statistically significant evidence of variance differences across peristalsis groups at the 5% level.
The Brown–Forsythe test yields p = 0.061, so there is no statistically significant evidence that the residual variance differs across peristalsis groups at the 5% level.

\qed

\problem
\textit{Proof}.

```{r}
quant_model = lm(respiratory_rate ~ temperature + pulse + packed_cell_volume + total_protein, data = horse)

summary(quant_model)

X = model.matrix(quant_model)
n = nrow(X)
p = ncol(X)
beta = coef(quant_model)
residuals = quant_model$residuals

```

```{r}
vcovOLS = summary(quant_model)$sigma^2 * summary(quant_model)$cov.unscaled
vcovOLS
confint(quant_model)

```

```{r}
#Heteroscedasticity-Consistent Estimators


library(sandwich)
vcovHC3 = vcovHC(quant_model, type = "HC3")
vcovHC3
```

```{r}
alpha = 0.05
HC3_CI = cbind(
  beta - qt(1 - alpha/2, n - p) * sqrt(diag(vcovHC3)),
  beta + qt(1 - alpha/2, n - p) * sqrt(diag(vcovHC3))
)

colnames(HC3_CI) = colnames(confint(quant_model))
HC3_CI
```

```{r}
library(lmtest)
confint(coeftest(quant_model, vcov. = vcovHC3))
```

```{r}
test = matrix(0, p, 4)
rownames(test) = names(beta)
colnames(test) = c("Estimate", "Std. Error", "t value", "Pr(>|t|)")

test[,1] = beta
test[,2] = sqrt(diag(vcovHC3))
test[,3] = test[,1] / test[,2]
test[,4] = 2 * pt(abs(test[,3]), n - p, lower.tail = FALSE)

test
```

```{r}

coeftest(quant_model, vcov. = vcovHC3)
```
The HC3 standard errors are larger for every coefficient. 
(Intercept) OLS SE: 52.6687 HC3 SE: 57.4952
temperature OLS SE:1.3922 HC3 SE:1.5417
pulse OLS SE:0.04231 HC3 SE:0.04909
PCV OLS SE: 0.10832 HC3 SE: 0.12214
total_protein OLS SE: 0.03676 HC3 SE:0.03822.

As the SEs increased, the HC3 t-values got smaller. 
(Intercept): -2.145 to -1.965
temperature: 2.442 to 2.205
pulse: 5.967 to 5.144
PCV: -0.640 to -0.568
total_protein: -1.011 to -0.972.

The p-values got larger. 
(Intercept)	0.0330 to 0.0506 (no longer < 0.05)	
temperature	0.0154 to 0.0284 (still < 0.05)	
pulse	8.82e-09 to 5.67e-07 (still extremely significant)	
PCV	0.5228 to 0.5709 (still NS)	
total_protein	0.3132 to 0.3319 (still NS)

The intercept is the only coefficient whose significance status changed. In the OLS (Intercept) is significant (p = 0.033), but HC3 shows it is only marginally significant (p = 0.0506).

The HC3 CIs are consistently wider:

(Intercept) OLS [-216.74, -9.22]; HC3 [-226.25, 0.29]. Which now covers zero.
temperature: OLS [0.656,6.142]; HC3 [0.362,6.43]
pulse: OLS [0.169,0.336]; HC3 [0.156,0.349]
PCV: OLS [-0.283, 0.144]; HC3 [-0.310, 0.171]
total_protein: OLS [-0.109, 0.0353]; HC3 [-0.112, 0.038]

The qualitative conclusions remained the same for temperature, pulse, PCV and total_protein. 
\qed


\problem
\textit{Proof}.
```{r}

```
\qed




