---
title: "SDS 4030 - Statistics for Data Science II"
subtitle: Final Project Code
author: 
output: pdf_document
fontsize: 12pt
header-includes:
  - \usepackage{hyperref}
  - \hypersetup{pdfpagemode={UseNone}, pdfstartview=}
  - \usepackage{amsthm}
  - \theoremstyle{definition}
  - \newtheorem{problem}{Problem}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(fig.align = "center", fig.height = 4, fig.width = 5)
knitr::opts_chunk$set(tidy.opts = list(width.cutoff = 68), tidy = TRUE)
knitr::opts_chunk$set(message = FALSE)
set.seed(0)
```

\problem
\textit{Proof}.
```{r}
# loading dataset Here

horse= read.table(file.choose(), header = TRUE)
head(horse)

```

```{r}
#fitted model w only quant variables

quant_model = lm(respiratory_rate ~ temperature + pulse + packed_cell_volume + total_protein, data = horse)
summary(quant_model)
```

```{r}
# qual, exact, large sample size -- lecture 21 reference


beta <- coef(quant_model)
X <- model.matrix(quant_model)
n <- nrow(X)
p <- ncol(X)
S2 <- sum(residuals(quant_model)^2) / (n - p) 


R <- matrix(c(
  0,  5,   0,    0, 350, 
  0,  0,  40,  100,   0 
), nrow = 2, byrow = TRUE)

r <- c(3, 3)
k <- length(r)

FT <- crossprod(R %*% beta - r, solve(R %*% solve(t(X) %*% X) %*% t(R), R %*% beta - r)) / (k * S2)
print(FT)

pvalue <- pf(FT, k, n - p, lower.tail = FALSE)
print(pvalue)

quant_reduced <- lm(respiratory_rate ~ I(temperature + total_protein + pulse + packed_cell_volume), data=horse)
summary(quant_reduced)


anova(quant_reduced, quant_model)


```

```{r}
# qual, asymptotic, large sample size -- lecture 21 reference

WT = k * FT
print(WT)
pvalue = pchisq(WT, k, lower.tail = FALSE)
print(pvalue)
anova(quant_reduced, quant_model, test = "Chisq")
```

The conclusions agree of both the exact and asymptotic tests agree with each other. Both had a p-value of 0.9043546, so we fail to reject the null hypothesis. The Wald statistic and Exact F statistic are related as 

\[W_n = \frac{\sigma}{S^2_{nQ1}} = k F\]

where F is the F statistic from the exact test. So we would expect the conclusions of both tests to agree with each other.


\qed

\problem
\textit{Proof}.

```{r}
#jackknife
betajack <- matrix(0, n, p)
for (i in 1:n) {
  jack <- lm(respiratory_rate ~ temperature + pulse + packed_cell_volume + total_protein,data = horse[-i, ])
  betajack[i, ] <- coef(jack)
}
betabar <- colMeans(betajack)
print(betabar)

SE <- apply(betajack, 2, sd) * sqrt((n - 1)^2 / n)
print(SE)

```

```{r}
#nonparametric bootstrap
nboot <- 10000
betaboot <- matrix(0, nboot, p)
set.seed(123)

for (k in 1:nboot) {
  indices <- sample(1:n, n, replace = TRUE)
  boot <- lm(respiratory_rate ~ temperature + pulse + packed_cell_volume + total_protein, data = horse[indices, ])
  betaboot[k, ] <- coef(boot)
}

betabar <- colMeans(betaboot)
print(betabar)

SE <- apply(betaboot, 2, sd)
print(SE)

```

```{r}
alpha <- 0.05

CInormal <- cbind(betabar - qnorm(1 - alpha / 2) * SE,
                  betabar + qnorm(1 - alpha / 2) * SE)
print(CInormal)

CIpercentile <- cbind(
  apply(betaboot, 2, quantile, probs = alpha / 2),
  apply(betaboot, 2, quantile, probs = 1 - alpha / 2)
)
print(CIpercentile)

beta_full <- coef(quant_model)
CIpivotal <- cbind(
  2 * beta_full - apply(betaboot, 2, quantile, probs = 1 - alpha / 2),
  2 * beta_full - apply(betaboot, 2, quantile, probs = alpha / 2)
)
print(CIpivotal)

```
```{r}
summary(quant_model)$coefficients[, "Std. Error"]

```
The coefficient estimates are essentially the same across methods (OLS / jackknife mean / bootstrap mean). Both resampling methods give very different estimates than OLS. The Jackknife SEs and bootstrap SEs are very similar. These SEs are much larger than OLS SEs. The two sets of intervals are very similar to one another, further supporting the stability of the bootstrap results. Compared to OLS confidence intervals, the bootstrap intervals are significantly wider.

\qed

\problem
\textit{Proof}.

```{r}
# Nonparametric test of staistical significant: Permutation F Test

f = summary(quant_model)$fstatistic[1]
print(f)
pvalue = pf(f, p, n - p - 1, lower.tail = FALSE)
print(pvalue)
nperm = 10000
fperm = numeric(nperm)
for (k in 1:nperm)
{
  Yperm = sample(Y)
  perm = lm(respiratory_rate ~ temperature + pulse + packed_cell_volume + total_protein, horse)
  fperm[k] = summary(perm)$fstatistic[1]
}
pvalue = (1 + sum(fperm > f)) / (1 + nperm)
print(pvalue)
```
The estimated F test p-value through the permutation method is equal to 1e-4. Since the pvalue is less than alpha, we can reject the null hypothesis. The pvalue under the theoretical null distribution is 6.4e-12, which is smaller than the pvalue from the permutation test. However, they both lead to the rejection of the null.

```{r}
# Nonparametric test of statistical significance for the effect of temperature
# Permutation t Test

t = summary(quant_model)$coefficients["temperature", 3]
print(t)
pvalue = 2 * pt(abs(t), n - p - 1, lower.tail = FALSE)
print(pvalue)
aux = lm(respiratory_rate ~ pulse + packed_cell_volume + total_protein, horse)
fitted = aux$fitted.values
residuals = aux$residuals
tperm = numeric(nperm)
for (k in 1:nperm)
{
Yperm = fitted + sample(residuals)
perm = lm(Yperm ~ temperature + pulse + packed_cell_volume + total_protein, horse)
tperm[k] = summary(perm)$coefficients["temperature", 3]
}
pvalue = (1 + sum(abs(tperm) > abs(t))) / (1 + nperm)
print(pvalue)

```
The estimated t test p-value through the permutation method is equal to 0.017. Since the pvalue is less than alpha, we can reject the null hypothesis. The pvalue under the theoretical null distribution is 0.015, which is very similar to the pvaqlue obtained from the permutation test.

\qed

\problem
\textit{Proof}.

```{r}
# Check for possible covariate transformations

par(mfrow = c(1, 5))
boxplot(horse$temperature, ylab = "Temperature", pch = 16)
boxplot(sqrt(horse$temperature), ylab = "Square Root of Temperature", pch = 16)
boxplot(log(horse$temperature), ylab = "Log of Temperature", pch = 16)
boxplot((horse$temperature)^2, ylab = "Square of Temperature", pch = 16)
boxplot((horse$temperature)^(-1), ylab = "Inverse of Temperature", pch = 16)

boxplot(horse$pulse, ylab = "Pulse", pch = 16)
boxplot(sqrt(horse$pulse), ylab = "Square Root of Pulse", pch = 16)
boxplot(log(horse$pulse), ylab = "Log of Pulse", pch = 16)
boxplot((horse$pulse)^2, ylab = "Square of Pulse", pch = 16)
boxplot((horse$pulse)^(-1), ylab = "Inverse of Pulse", pch = 16)

boxplot(horse$total_protein, ylab = "Total Protein", pch = 16)
boxplot(sqrt(horse$total_protein), ylab = "Square Root of Total Protein", pch = 16)
boxplot(log(horse$total_protein), ylab = "Log of Total Protein", pch = 16)
boxplot((horse$total_protein)^2, ylab = "Square of Total Protein", pch = 16)
boxplot((horse$total_protein)^(-1), ylab = "Inverse of Total Protein", pch = 16)

boxplot(horse$packed_cell_volume, ylab = "Packed Cell Volume", pch = 16)
boxplot(sqrt(horse$packed_cell_volume), ylab = "Square Root of Packed Cell Volume", pch = 16)
boxplot(log(horse$packed_cell_volume), ylab = "Log of Packed Cell Volume", pch = 16)
boxplot((horse$packed_cell_volume)^2, ylab = "Square of Packed Cell Volume", pch = 16)
boxplot((horse$packed_cell_volume)^(-1), ylab = "Inverse of Packed Cell Volume", pch = 16)
```
Pulse and total protein seem skewed, so need transformations.
For pulse, the best transformation is inverse. For total protein, the best transformation is a log transformation. A log transformation is reasonable for packed cell volume to decrease outliers and make for a more symmetric distribution.

```{r}
fit = lm(respiratory_rate ~ temperature + (1/pulse) + log(packed_cell_volume) + log(total_protein), horse)
plot(fit$fitted.values, rstandard(fit), xlab = "Fitted Values",
ylab = "Standardized Residuals", pch = 16)
abline(h = 0, lty = 2, lwd = 2)
```
Since constant variance is violated, a transformation should be applied to the response variable.

```{r} 
#Box-Cox transformation
horse$inv_pulse = 1 / horse$pulse
Y = horse$respiratory_rate

lambda = seq(-2, 2, 0.01)
loglik = numeric(401)
for (i in 1:401)
{
  if (lambda[i] == 0)
  {
    Ypower = log(Y)
  }
  else
  {
    Ypower = (Y^lambda[i] - 1) / lambda[i]
  }
  power = lm(Ypower ~ temperature + inv_pulse + log(packed_cell_volume) + log(total_protein), horse)
  loglik[i] = logLik(power)[1] + (lambda[i] - 1) * sum(log(Y))
}

CI = range(lambda[loglik > max(loglik) - qchisq(alpha, 1, lower.tail = FALSE) / 2])
print(CI)
```
Since the CI is close to covering 0, a log transformation is reasonable.

```{r}

power = lm(log(respiratory_rate) ~ temperature + inv_pulse + log(packed_cell_volume) + log(total_protein), horse)
plot(power$fitted.values, rstandard(power), main = "Log-
Transformation", xlab = "Fitted Values", ylab = "
Standardized Residuals", pch = 16)
abline(h = 0, lty = 2, lwd = 2)

library(qqconf)
qq_conf_plot(rstandard(power), points_params = list(pch = 16, cex = 0.5))

```
All standardized residuals fall within the band, except one. This is enough to reject normality.

\qed

\problem
\textit{Proof}.
```{r}

#Best subset selection with valid post-selection inference 

library(leaps)

#New horse data frame with 5 transformed variables
horse_transformed = data.frame(
  Y = log(horse$respiratory_rate),
  temperature = horse$temperature,
  inv_pulse = 1 / horse$pulse,
  log_pcv  = log(horse$packed_cell_volume),
  log_tp   = log(horse$total_protein)
)

Y = horse_transformed$Y
X =  horse_transformed[, -1]      
n = nrow(X)
P = ncol(X)


#Set.seed for consistency
set.seed(111)

train = sample(n, 0.6 * n)
valid = sample(setdiff(1:n, train), 0.2 * n)
test  =  setdiff(1:n, c(train, valid))
MSPE = numeric(P)

full = lm(Y ~ ., horse_transformed, train)
sub = regsubsets(formula(full), horse_transformed[train, ], nvmax = P)
summary(sub)$which

for (p in 1:P)
{
  fit = lm(Y ~ ., X[, summary(sub)$which[p, -1], drop = FALSE], train)
  preds = predict(fit, X[valid, summary(sub)$which[p, -1], drop = FALSE])
  MSPE[p] = mean((Y[valid] - preds)^2)
}

plot(MSPE, type="b", pch=16, lwd=4, ylab="Validation MSPE")

p =  which.min(MSPE)
final_model = lm(Y ~ ., X[, summary(sub)$which[p, -1], drop = FALSE], test)
summary(final_model)


```

Based on the validation MSPE curve, the full (4 covariates) model provides the most accurate predictor for respiratory rate. Among all of the predictors, inverse pulse (inv_pulse) is the only statistically significant variable on the test set and shows that horses with higher pulse rates tend to have higher respiratory rates. The remaining predictors (temperature, log(packed cell volume), and log(total protein)) are not individually statistically significant.

\qed

\problem
\textit{Proof}.
```{r}
#Investigate collinearity 
library(faraway)
cor(horse_transformed)
plot(horse_transformed, pch = 16, cex = 0.25)

cat("VIF:", vif(full))

```

Based on the VIFs and correlation plots, the transformed quantitative variables aren't very collinear. 
```{r}
#Ridge Regression 
library(glmnet)
train_boot = c(train, valid)
X = as.matrix(X)
cv.ridge = cv.glmnet(X[train_boot,], Y[train_boot], nfolds = 6, alpha = 0)
ridge = glmnet(X[test,], Y[test], alpha = 0, lambda = cv.ridge$lambda.min)
print(ridge$beta)

#Bootstrap
nboot = 10000
beta = matrix(0, nboot, P)
colnames(beta) = colnames(X)
for (i in 1:nboot)
{
  ind = sample(test, replace = TRUE)
  beta[i,] = glmnet(X[ind,], Y[ind], alpha = 0, lambda = cv.ridge$lambda.min)$beta[,]
}
SE = apply(beta, 2, sd)
cat("SE after Bootstrap:", SE, "\n")


ols = lm(Y ~ ., horse_transformed[train, ])
cat("Summary of OLS coefficients for comparison:","\n")
summary(ols)$coefficients


```

The ridge regression coefficients all shrunk closer to zero in comparison to the OLS coefficients. The bootstrap results SEs are slightly higher than the SEs calculated using OLS. 

\qed

\problem
\textit{Proof}.

```{r}
#Lasso Regression
cv.lasso = cv.glmnet(X[train,], Y[train], alpha = 1)
lasso = glmnet(X[test,], Y[test], nfolds = 6, alpha = 1, lambda
= cv.lasso$lambda.min)
print(lasso$beta)

for (i in 1:nboot)
{
  ind = sample(test, replace = TRUE)
  beta[i,] = glmnet(X[ind,], Y[ind], alpha = 1, lambda = cv.lasso$lambda.min)$beta[,]
}
SE = apply(beta, 2, sd)
cat("SE after Bootstrap:", SE, "\n")
```
While some of the estimated lasso coefficients become extremely close to zero, others stay the same. Because the quantitative variables aren't collinear, the estimated SEs of the lasso coefficients are greater than the SEs estimated by the OLS. 
\qed

\problem
\textit{Proof}.

```{r}
#Investigate the interaction effects between pain and all available quantitative covariates on the response variable.

Y = log(horse$respiratory_rate)
inv_pulse = 1 / horse$pulse
log_pcv  = log(horse$packed_cell_volume)
log_tp   = log(horse$total_protein)
horse$pain = as.factor(horse$pain)


full_interact = lm( Y ~ pain*(temperature + inv_pulse + log_pcv + log_tp), horse)
summary(full_interact)

anova(full_interact)
```
```{r}
fit_interact = update(full_interact, . ~ . - pain:log_pcv)
summary(fit_interact)
anova(fit_interact)
```

```{r}
fit_interact = update(fit_interact, . ~ . - pain:log_pcv)
summary(fit_interact)
anova(fit_interact)
```
```{r}
fit_interact = update(fit_interact, . ~ . - log_pcv)
summary(fit_interact)
anova(fit_interact)
```

```{r}
fit_interact = update(fit_interact, . ~ . - log_tp)
summary(fit_interact)
anova(fit_interact)
```
```{r}
fit_interact = update(fit_interact, . ~ . - paincontinuous:temperature)
summary(fit_interact)
anova(fit_interact)
```


Final Model Equation: 
\[
\begin{aligned}
Y_i
&= \beta_0
+ \beta_1\,\text{pain}_{i,\text{continuous}}
+ \beta_2\,\text{pain}_{i,\text{depressed}}
+ \beta_3\,\text{pain}_{i,\text{mild}}
+ \beta_4\,\text{pain}_{i,\text{severe}} \\
&\quad + \beta_5\,\text{temperature}_i
+ \beta_6\,\text{inv\_pulse}_i \\
&\quad + \gamma_1\,\text{pain}_{i,\text{continuous}}\cdot\text{temperature}_i
+ \gamma_2\,\text{pain}_{i,\text{depressed}}\cdot\text{temperature}_i
+ \gamma_3\,\text{pain}_{i,\text{mild}}\cdot\text{temperature}_i
+ \gamma_4\,\text{pain}_{i,\text{severe}}\cdot\text{temperature}_i \\
&\quad + \delta_1\,\text{pain}_{i,\text{continuous}}\cdot\text{inv\_pulse}_i
+ \delta_2\,\text{pain}_{i,\text{depressed}}\cdot\text{inv\_pulse}_i
+ \delta_3\,\text{pain}_{i,\text{mild}}\cdot\text{inv\_pulse}_i
+ \delta_4\,\text{pain}_{i,\text{severe}}\cdot\text{inv\_pulse}_i
+ \varepsilon_i.
\end{aligned}
\]


Interpretation of each covariate in the final model: 

- paincontinuous = –12.1040
Holding other predictors constant, a horse classified as continuous pain is expected to have a log respiratory rate about 12.10 units lower than a horse with no pain.

- paindepressed = –17.5964
Holding other predictors constant, a depressed horse has a log respiratory rate about 17.60 units lower than a horse with no pain.

- painmild = –10.8063
Holding other predictors constant, a horse with mild pain has a log respiratory rate about 10.81 units lower than a no-pain horse.

- painsevere = –13.0211
Holding other predictors constant, a horse with severe pain has a log respiratory rate about 13.02 units lower than a no-pain horse.

-temperature = –0.2557
For a horse in the no-pain category with a fixed pulse, a 1°C increase in temperature is associated with a 0.256-unit decrease in log respiratory rate. Temperature is significant overall (ANOVA p = 0.0227), but its slope for no-pain horses is not individually significant.

- inv_pulse = –64.7889
Because inv_pulse = 1/pulse, a decrease in inv_pulse corresponds to a higher pulse rate. For a horse in the no-pain category, holding temperature constant, a one-unit increase in inv_pulse is associated with a 64.79-unit decrease in log respiratory rate.

- paindepressed:temperature = 0.4473
For depressed horses, the effect of temperature on log respiratory rate is 0.447 units more positive per degree than for no-pain horses.This is the only significant interaction in the model (p = 0.0151).
