---
title: "SDS 4030 - Statistics for Data Science II"
subtitle: Final Project Code
author: 
output: pdf_document
fontsize: 12pt
header-includes:
  - \usepackage{hyperref}
  - \hypersetup{pdfpagemode={UseNone}, pdfstartview=}
  - \usepackage{amsthm}
  - \theoremstyle{definition}
  - \newtheorem{problem}{Problem}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(fig.align = "center", fig.height = 4, fig.width = 5)
knitr::opts_chunk$set(tidy.opts = list(width.cutoff = 68), tidy = TRUE)
knitr::opts_chunk$set(message = FALSE)
set.seed(0)
```

\problem
\textit{Proof}.
```{r}
# loading dataset Here

horse= read.table("horse.txt", header = TRUE)
head(horse)

```

```{r}
#starter EDA on resp rate
library(ggplot2)

resp_summary <- summary(horse$respiratory_rate)
resp_summary

mean_resp <- mean(horse$respiratory_rate, na.rm = TRUE)
median_resp <- median(horse$respiratory_rate, na.rm = TRUE)
min_resp <- min(horse$respiratory_rate, na.rm = TRUE)
max_resp <- max(horse$respiratory_rate, na.rm = TRUE)

cat("Mean:", mean_resp, "\n")
cat("Median:", median_resp, "\n")
cat("Min:", min_resp, "\n")
cat("Max:", max_resp, "\n")

ggplot(horse, aes(x = respiratory_rate)) +
  geom_histogram(bins = 15, fill = "gray", color = "black") +
  geom_vline(xintercept = 8, linetype = "dashed", color = "red", size=1) +
  geom_vline(xintercept = 16, linetype = "dashed", color = "red", size=1) +
  annotate("rect", xmin=8, xmax=16, ymin=0, ymax=Inf, alpha=0.1, fill="red") +
  labs(title = "Respiratory Rate Distribution with Normal Range (8-16)",
       x = "Respiratory Rate (breaths/min)",
       y = "Count")


```

```{r}
#correlation matrix for now

library(corrplot)

numeric_vars <- c("respiratory_rate", "pulse", "temperature", "packed_cell_volume", "total_protein")
corr_matrix <- cor(horse[, numeric_vars], use = "complete.obs")

corrplot(corr_matrix, method = "color", addCoef.col = "black", tl.cex = 0.8)

```

```{r}
#fitted model w only quant variables

quant_model = lm(respiratory_rate ~ temperature + pulse + packed_cell_volume + total_protein, data = horse)
summary(quant_model)
```

```{r}
# qual, exact, large sample size -- lecture 21 reference


beta <- coef(quant_model)
X <- model.matrix(quant_model)
Y = horse$respiratory_rate
n <- nrow(X)
p <- ncol(X)
S2 <- sum(residuals(quant_model)^2) / (n - p) 


R <- matrix(c(
  0,  5,   0,    0, 350, 
  0,  0,  40,  100,   0 
), nrow = 2, byrow = TRUE)

r <- c(3, 3)
k <- length(r)

FT <- crossprod(R %*% beta - r, solve(R %*% solve(t(X) %*% X) %*% t(R), R %*% beta - r)) / (k * S2)
print(FT)

pvalue <- pf(FT, k, n - p, lower.tail = FALSE)
print(pvalue)

quant_reduced <- lm(respiratory_rate ~ I(temperature + total_protein + pulse + packed_cell_volume), data=horse)
summary(quant_reduced)


anova(quant_reduced, quant_model)


```

```{r}
# qual, asymptotic, large sample size -- lecture 21 reference

WT = k * FT
print(WT)
pvalue = pchisq(WT, k, lower.tail = FALSE)
print(pvalue)
anova(quant_reduced, quant_model, test = "Chisq")
```

The conclusions agree of both the exact and asymptotic tests agree with each other. Both had a p-value of 0.9043546, so we fail to reject the null hypothesis. The Wald statistic and Exact F statistic are related as 

\[W_n = \frac{\sigma}{S^2_{nQ1}} = k F\]

where F is the F statistic from the exact test. So we would expect the conclusions of both tests to agree with each other.


\qed

\problem
\textit{Proof}.

```{r}
#jackknife
betajack <- matrix(0, n, p)
for (i in 1:n) {
  jack <- lm(respiratory_rate ~ temperature + pulse + packed_cell_volume + total_protein,data = horse[-i, ])
  betajack[i, ] <- coef(jack)
}
betabar <- colMeans(betajack)
print(betabar)

SE <- apply(betajack, 2, sd) * sqrt((n - 1)^2 / n)
print(SE)

```

```{r}
#nonparametric bootstrap
nboot <- 10000
betaboot <- matrix(0, nboot, p)
set.seed(123)

for (k in 1:nboot) {
  indices <- sample(1:n, n, replace = TRUE)
  boot <- lm(respiratory_rate ~ temperature + pulse + packed_cell_volume + total_protein, data = horse[indices, ])
  betaboot[k, ] <- coef(boot)
}

betabar <- colMeans(betaboot)
print(betabar)

SE <- apply(betaboot, 2, sd)
print(SE)

```

```{r}
alpha <- 0.05

CInormal <- cbind(betabar - qnorm(1 - alpha / 2) * SE,
                  betabar + qnorm(1 - alpha / 2) * SE)
print(CInormal)

CIpercentile <- cbind(
  apply(betaboot, 2, quantile, probs = alpha / 2),
  apply(betaboot, 2, quantile, probs = 1 - alpha / 2)
)
print(CIpercentile)

beta_full <- coef(quant_model)
CIpivotal <- cbind(
  2 * beta_full - apply(betaboot, 2, quantile, probs = 1 - alpha / 2),
  2 * beta_full - apply(betaboot, 2, quantile, probs = alpha / 2)
)
print(CIpivotal)

```
```{r}
summary(quant_model)$coefficients[, "Std. Error"]

```
The coefficient estimates are essentially the same across methods (OLS / jackknife mean / bootstrap mean). Both resampling methods give very different estimates than OLS. The Jackknife SEs and bootstrap SEs are very similar. These SEs are much larger than OLS SEs. The two sets of intervals are very similar to one another, further supporting the stability of the bootstrap results. Compared to OLS confidence intervals, the bootstrap intervals are significantly wider.

\qed

\problem
\textit{Proof}.

```{r}
# Nonparametric test of staistical significant: Permutation F Test

f = summary(quant_model)$fstatistic[1]
print(f)
pvalue = pf(f, p, n - p - 1, lower.tail = FALSE)
print(pvalue)
nperm = 10000
fperm = numeric(nperm)
for (k in 1:nperm)
{
  Yperm = sample(Y)
  perm = lm(respiratory_rate ~ temperature + pulse + packed_cell_volume + total_protein, horse)
  fperm[k] = summary(perm)$fstatistic[1]
}
pvalue = (1 + sum(fperm > f)) / (1 + nperm)
print(pvalue)
```
The estimated F test p-value through the permutation method is equal to 1e-4. Since the pvalue is less than alpha, we can reject the null hypothesis. The pvalue under the theoretical null distribution is 6.4e-12, which is smaller than the pvalue from the permutation test. However, they both lead to the rejection of the null.

```{r}
# Nonparametric test of statistical significance for the effect of temperature
# Permutation t Test

t = summary(quant_model)$coefficients["temperature", 3]
print(t)
pvalue = 2 * pt(abs(t), n - p - 1, lower.tail = FALSE)
print(pvalue)
aux = lm(respiratory_rate ~ pulse + packed_cell_volume + total_protein, horse)
fitted = aux$fitted.values
residuals = aux$residuals
tperm = numeric(nperm)
for (k in 1:nperm)
{
Yperm = fitted + sample(residuals)
perm = lm(Yperm ~ temperature + pulse + packed_cell_volume + total_protein, horse)
tperm[k] = summary(perm)$coefficients["temperature", 3]
}
pvalue = (1 + sum(abs(tperm) > abs(t))) / (1 + nperm)
print(pvalue)

```
The estimated t test p-value through the permutation method is equal to 0.017. Since the pvalue is less than alpha, we can reject the null hypothesis. The pvalue under the theoretical null distribution is 0.015, which is very similar to the pvaqlue obtained from the permutation test.

\qed

\problem
\textit{Proof}.

```{r}
# Check for possible covariate transformations

par(mfrow = c(1, 5))
boxplot(horse$temperature, ylab = "Temperature", pch = 16)
boxplot(sqrt(horse$temperature), ylab = "Square Root of Temperature", pch = 16)
boxplot(log(horse$temperature), ylab = "Log of Temperature", pch = 16)
boxplot((horse$temperature)^2, ylab = "Square of Temperature", pch = 16)
boxplot((horse$temperature)^(-1), ylab = "Inverse of Temperature", pch = 16)

boxplot(horse$pulse, ylab = "Pulse", pch = 16)
boxplot(sqrt(horse$pulse), ylab = "Square Root of Pulse", pch = 16)
boxplot(log(horse$pulse), ylab = "Log of Pulse", pch = 16)
boxplot((horse$pulse)^2, ylab = "Square of Pulse", pch = 16)
boxplot((horse$pulse)^(-1), ylab = "Inverse of Pulse", pch = 16)

boxplot(horse$total_protein, ylab = "Total Protein", pch = 16)
boxplot(sqrt(horse$total_protein), ylab = "Square Root of Total Protein", pch = 16)
boxplot(log(horse$total_protein), ylab = "Log of Total Protein", pch = 16)
boxplot((horse$total_protein)^2, ylab = "Square of Total Protein", pch = 16)
boxplot((horse$total_protein)^(-1), ylab = "Inverse of Total Protein", pch = 16)

boxplot(horse$packed_cell_volume, ylab = "Packed Cell Volume", pch = 16)
boxplot(sqrt(horse$packed_cell_volume), ylab = "Square Root of Packed Cell Volume", pch = 16)
boxplot(log(horse$packed_cell_volume), ylab = "Log of Packed Cell Volume", pch = 16)
boxplot((horse$packed_cell_volume)^2, ylab = "Square of Packed Cell Volume", pch = 16)
boxplot((horse$packed_cell_volume)^(-1), ylab = "Inverse of Packed Cell Volume", pch = 16)
```
Pulse and total protein seem skewed, so need transformations.
For pulse, the best transformation is inverse. For total protein, the best transformation is a log transformation. A log transformation is reasonable for packed cell volume to decrease outliers and make for a more symmetric distribution.

```{r}
fit = lm(respiratory_rate ~ temperature + (1/pulse) + log(packed_cell_volume) + log(total_protein), horse)
plot(fit$fitted.values, rstandard(fit), xlab = "Fitted Values",
ylab = "Standardized Residuals", pch = 16)
abline(h = 0, lty = 2, lwd = 2)
```
Since constant variance is violated, a transformation should be applied to the response variable.

```{r} 
#Box-Cox transformation
horse$inv_pulse = 1 / horse$pulse
Y = horse$respiratory_rate

lambda = seq(-2, 2, 0.01)
loglik = numeric(401)
for (i in 1:401)
{
  if (lambda[i] == 0)
  {
    Ypower = log(Y)
  }
  else
  {
    Ypower = (Y^lambda[i] - 1) / lambda[i]
  }
  power = lm(Ypower ~ temperature + inv_pulse + log(packed_cell_volume) + log(total_protein), horse)
  loglik[i] = logLik(power)[1] + (lambda[i] - 1) * sum(log(Y))
}

CI = range(lambda[loglik > max(loglik) - qchisq(alpha, 1, lower.tail = FALSE) / 2])
print(CI)
```
Since the CI is close to covering 0, a log transformation is reasonable.

```{r}

power = lm(log(respiratory_rate) ~ temperature + inv_pulse + log(packed_cell_volume) + log(total_protein), horse)
plot(power$fitted.values, rstandard(power), main = "Log-
Transformation", xlab = "Fitted Values", ylab = "
Standardized Residuals", pch = 16)
abline(h = 0, lty = 2, lwd = 2)

library(qqconf)
qq_conf_plot(rstandard(power), points_params = list(pch = 16, cex = 0.5))

```
All standardized residuals fall within the band, except one. This is enough to reject normality.

\qed

\problem
\textit{Proof}.
```{r}

#Best subset selection with valid post-selection inference 

library(leaps)

#New horse data frame with 5 transformed variables
horse_transformed = data.frame(
  Y = log(horse$respiratory_rate),
  temperature = horse$temperature,
  inv_pulse = 1 / horse$pulse,
  log_pcv  = log(horse$packed_cell_volume),
  log_tp   = log(horse$total_protein)
)

Y = horse_transformed$Y
X =  horse_transformed[, -1]      
n = nrow(X)
P = ncol(X)

train = sample(n, 0.6 * n)
valid = sample(setdiff(1:n, train), 0.2 * n)
test  =  setdiff(1:n, c(train, valid))
MSPE = numeric(P)

full = lm(Y ~ ., horse_transformed, train)
sub = regsubsets(formula(full), horse_transformed[train, ], nvmax = P)
summary(sub)$which

for (p in 1:P)
{
  fit = lm(Y ~ ., X[, summary(sub)$which[p, -1], drop = FALSE], train)
  preds = predict(fit, X[valid, summary(sub)$which[p, -1], drop = FALSE])
  MSPE[p] = mean((Y[valid] - preds)^2)
}

plot(MSPE, type="b", pch=16, lwd=4, ylab="Validation MSPE")

p =  which.min(MSPE)
final_model = lm(Y ~ ., X[, summary(sub)$which[p, -1], drop = FALSE], test)
summary(final_model)


```

Based on the validation MSPE curve, the full (4 covariates) model provides the most accurate predictor for respiratory rate. Among all of the predictors, inverse pulse (inv_pulse) is the only statistically significant variable on the test set and shows that horses with higher pulse rates tend to have higher respiratory rates. The remaining predictors (temperature, log(packed cell volume), and log(total protein)) are not individually statistically significant.

\qed

\problem
\textit{Proof}.
```{r}
#Investigate collinearity 
library(faraway)
cor(horse_transformed)
plot(horse_transformed, pch = 16, cex = 0.25)

cat("VIF:", vif(full))

```

Based on the VIFs and correlation plots, the transformed quantitative variables aren't very collinear. 

```{r}
#Ridge Regression 
library(glmnet)
train_boot = c(train, valid)
X = as.matrix(X)
cv.ridge = cv.glmnet(X[train_boot,], Y[train_boot], nfolds = 6, alpha = 0)
ridge = glmnet(X[test,], Y[test], alpha = 0, lambda = cv.ridge$lambda.min)
print(ridge$beta)

#Bootstrap
nboot = 10000
beta = matrix(0, nboot, P)
colnames(beta) = colnames(X)
for (i in 1:nboot)
{
  ind = sample(test, replace = TRUE)
  beta[i,] = glmnet(X[ind,], Y[ind], alpha = 0, lambda = cv.ridge$lambda.min)$beta[,]
}
SE = apply(beta, 2, sd)
cat("SE after Bootstrap:", SE, "\n")


ols = lm(Y ~ ., horse_transformed[train, ])
cat("Summary of OLS coefficients for comparison:","\n")
summary(ols)$coefficients


```

The ridge regression coefficients all shrunk closer to zero in comparison to the OLS coefficients. The bootstrap results SEs are slightly higher than the SEs calculated using OLS. 

\qed

\problem
\textit{Proof}.

```{r}
#Lasso Regression
cv.lasso = cv.glmnet(X[train,], Y[train], alpha = 1)
lasso = glmnet(X[test,], Y[test], nfolds = 6, alpha = 1, lambda
= cv.lasso$lambda.min)
print(lasso$beta)

for (i in 1:nboot)
{
  ind = sample(test, replace = TRUE)
  beta[i,] = glmnet(X[ind,], Y[ind], alpha = 1, lambda = cv.lasso$lambda.min)$beta[,]
}
SE = apply(beta, 2, sd)
cat("SE after Bootstrap:", SE, "\n")
```
While some of the estimated lasso coefficients become extremely close to zero, others stay the same. Because the quantitative variables aren't collinear, the estimated SEs of the lasso coefficients are greater than the SEs estimated by the OLS. 
\qed

\problem
\textit{Proof}.

```{r}
#Investigate the interaction effects between pain and all available quantitative covariates on the response variable.

#Factor qualitative variable
horse$pain = as.factor(horse$pain)

Y = log(horse$respiratory_rate)
inv_pulse = 1 / horse$pulse
log_pcv  = log(horse$packed_cell_volume)
log_tp   = log(horse$total_protein)
horse$pain = as.factor(horse$pain)


fit_interact = lm( Y ~ pain*(temperature + inv_pulse + log_pcv + log_tp), horse)
anova(fit_interact)

```
```{r}
fit_interact = update(fit_interact, . ~ . - pain:log_tp)
anova(fit_interact)
```

```{r}
fit_interact = update(fit_interact, . ~ . - pain:inv_pulse)
anova(fit_interact)
```

```{r}
fit_interact = update(fit_interact, . ~ . - pain:log_pcv)
anova(fit_interact)
```

```{r}
fit_interact = update(fit_interact, . ~ . - log_pcv)
anova(fit_interact)
```
```{r}
fit_interact = update(fit_interact, . ~ . - pain:temperature)
anova(fit_interact)
```
```{r}
fit_interact = update(fit_interact, . ~ . - log_tp)
anova(fit_interact)
```
```{r}
summary(fit_interact)
```
Final Model: 
\[
Y = \beta_0 
+ \beta_1 \,\text{pain}_{\text{continuous}}
+ \beta_2 \,\text{pain}_{\text{depressed}}
+ \beta_3 \,\text{pain}_{\text{mild}}
+ \beta_4 \,\text{pain}_{\text{severe}}
+ \beta_5 \,\text{temperature}
+ \beta_6 \,\text{inv\_pulse}
+ \varepsilon
\]



Interpretation of each covariate in the final model: 

pain (continuous): Horses in the continuous pain category have an expected log-respiratory-rate that is 0.171 units higher than horses in the reference pain category, holding temperature and inverse pulse constant.

pain (depressed): Horses rated as depressed have an expected log-respiratory-rate that is 0.126 units higher than horses in the reference pain category, adjusting for temperature and inverse pulse.

pain (mild): Horses with mild pain have an expected log-respiratory-rate that is 0.076 units higher than the reference category, controlling for temperature and inverse pulse.

pain (severe): Horses with severe pain have an expected log-respiratory-rate that is 0.090 units higher than horses in the reference category, holding temperature and inverse pulse constant.

temperature: For each 1-degree increase in temperature, the expected log-respiratory-rate increases by 0.0749 units, controlling for pain level and inverse pulse.

inverse pulse (inv_pulse): For each 1-unit increase in inverse pulse (which corresponds to a decrease in pulse rate), the expected log-respiratory-rate decreases by 36.61 units, holding pain level and temperature constant.

\qed

\problem
\textit{Proof}.
```{r}
#Mixed-effects Model

Y = horse$respiratory_rate
horse$pain = factor(horse$pain)
horse$surgical_lesion = factor(horse$surgical_lesion)
horse$peristalsis = factor(horse$peristalsis)
horse$abdominal_distension = factor(horse$abdominal_distension)
horse$peripheral_pulse = factor(horse$peripheral_pulse)
horse$temperature_of_extremities = factor(horse$temperature_of_extremities)
horse$surgery <- factor(horse$surgery)
horse$outcome <- factor(horse$outcome)
horse$age <- factor(horse$age)

```

```{r}
#chosen random effects

# Used transformed variables from problem 8 for reference

library(lme4)
mixed_eff_mod = lmer(
  Y ~ temperature + inv_pulse + log_pcv + log_tp + outcome + surgery + abdominal_distension + pain + 
(1 | temperature_of_extremities) +(1 | peripheral_pulse) + (1 | age) + (1 |surgical_lesion ) + (1 | peristalsis),
  data = horse)

```

```{r}
sigma = data.frame(summary(mixed_eff_mod)$varcor)$sdcor
names(sigma) = data.frame(summary(mixed_eff_mod)$varcor)$grp
```

```{r}
ICC = sigma^2 / sum(sigma^2)
print(ICC)
```

```{r}
summary(mixed_eff_mod, correlation = FALSE)
anova(mixed_eff_mod)
```
```{r}
#Box Cox
Y = horse$respiratory_rate

lambda = seq(-2, 2, 0.01)
loglik = numeric(401)
for (i in 1:401)
{
  if (lambda[i] == 0)
  {
    Ypower = log(Y)
  }
  else
  {
    Ypower = (Y^lambda[i] - 1) / lambda[i]
  }
  power = lmer(Ypower ~ temperature + inv_pulse + log_pcv + log_tp + outcome + surgery + abdominal_distension + pain + 
(1 | temperature_of_extremities) +(1 | peripheral_pulse) + (1 | age) + (1 |surgical_lesion ) + (1 | peristalsis), horse)
  loglik[i] = logLik(power)[1] + (lambda[i] - 1) * sum(log(Y))
}

```

```{r}
alpha = 0.05
CI = range(lambda[loglik > max(loglik) - qchisq(alpha, 1, lower.tail = FALSE) / 2])
print(CI)
plot(lambda, loglik, "l", xlab = expression(lambda), ylab = "
Profile Log-Likelihood", lwd = 4)
abline(h = max(loglik) - qchisq(0.95, 1) / 2, lty = 2, lwd = 4)
abline(v = CI, lty = 2, lwd = 4)
```


The covariates chosen as random effects are age, peripheral_pulse, peristalsis, and temperature_of_extremities, since these were the categorical variables with the largest nonzero variance components in the mixed-effects model. All categorical variables with exactly zero variance components were treated as fixed effects, as instructed.

A suitable transformation for our response variable is.

The total variation in your response variable explained by each random effect is as follows: abdominal_distension (variance = 0.02668, 2.67%) and temperature_of_extremities (variance = 0.00652, 0.65%), and surgery (variance = 0.005579, 0.56%).


Another interpretation for each ICC:
ICC(abdominal_distension) = 0.0936
Two horses with the same abdominal distension level have an expected correlation of about 9.4% in their log respiratory rates. This points too a moderate grouping effect of adominal distension level on respiratory rate.

ICC(temperature_of_extremities) = 0.0229
Two horses with the same extremity temperature category have an expected correlation of about 2.3% in their log respiratory rates. This points to a very small grouping effect of extremity temperature on respiratory rate.

ICC(surgery) = 0.01956768
Two horses that underwent the same type of surgery have an expected correlation of about 2.0% in their log respiratory rates. This points to a very small grouping effect of surgery on respiratory rate.



\qed

\problem
\textit{Proof}.
```{r}
base_model <- lm(Y ~ temperature + inv_pulse + log_pcv + log_tp, data = horse)
print(summary(base_model))
```

```{r}
Y = log(horse$respiratory_rate)
X = model.matrix(base_model)
n = dim(X)[1]
p = dim(X)[2] - 1
residuals = base_model$residuals
plot(rstandard(base_model) ~ temperature + inv_pulse + log_pcv + log_tp, horse, ylab = "Standardized
Residuals", pch = 16)
abline(h = 0, lty = 2, lwd = 2)
```
```{r}
#Studentized Breusch-Pagan Test
Yaux = residuals(base_model)^2
aux = lm(Yaux ~ temperature + inv_pulse + log_pcv + log_tp, horse)
BP = n * summary(aux)$r.squared
print(BP)
pvalue = pchisq(BP, p, lower.tail = FALSE)
print(pvalue)
library(lmtest)
bptest(base_model)
```

```{r}
#Brown-Forsythe Test -- pain

library(car)
leveneTest(Y ~ pain, data = horse)
```

```{r}
#Brown-Forsythe Test -- outcome
library(car)
leveneTest(Y ~ outcome, data = horse)

```
```{r}
#Brown-Forsythe Test -- temp of extremities

library(car)
leveneTest(Y ~ temperature_of_extremities, data = horse)

```

```{r}
#Brown-Forsythe Test -- peripheral pulse
library(car)
leveneTest(Y ~ peripheral_pulse, data = horse)
```

```{r}
#Brown-Forsythe Test -- abdominal distension

library(car)
leveneTest(Y ~ abdominal_distension, data = horse)

```

```{r}
#Brown-Forsythe Test -- peristalsis

library(car)
leveneTest(Y ~ peristalsis, data = horse)
```
For the quantitative variables, a studentized Breusch–Pagan test was conducted. Because the p-value is well above 0.05, we fail to reject the null hypothesis of constant variance. So here is no evidence of heteroscedasticity, and the residual variance does not appear to depend on any of the included quantitative predictors.


For the qualitative varibles, a Brown–Forsythe tests was conducted. 

The Brown–Forsythe test yields p = 0.0006161, providing statistically significant evidence that the residual variance differs across pain groups (heteroscedasticity).

The Brown–Forsythe test yields p = 0.2814, so there is no statistically significant evidence of differing variances across outcome groups (homoscedasticity).

The Brown–Forsythe test yields p = 0.4136, so there is no statistically significant evidence that the residual variance differs across peripheral-pulse categories.

The Brown–Forsythe test yields p = 0.0444, indicating statistically significant evidence that residual variances differ across abdominal-distension groups (heteroscedasticity).

The Brown–Forsythe test yields p = 0.061, so there is no statistically significant evidence of variance differences across peristalsis groups at the 5% level.
The Brown–Forsythe test yields p = 0.061, so there is no statistically significant evidence that the residual variance differs across peristalsis groups at the 5% level.

\qed

\problem
\textit{Proof}.

```{r}

Y = log(horse$respiratory_rate)
quant_model = lm(Y ~ temperature + inv_pulse + log_pcv + log_tp, horse)

ols_summary = summary(quant_model)

X = model.matrix(quant_model)
n = nrow(X)
p = ncol(X)
beta = coef(quant_model)
residuals = quant_model$residuals

```

```{r}
vcovOLS = summary(quant_model)$sigma^2 * summary(quant_model)$cov.unscaled
vcovOLS
confint(quant_model)

```

```{r}
#Heteroscedasticity-Consistent Estimators


library(sandwich)
vcovHC3 = vcovHC(quant_model, type = "HC3")
vcovHC3
```

```{r}
alpha = 0.05
HC3_CI = cbind(
  beta - qt(1 - alpha/2, n - p) * sqrt(diag(vcovHC3)),
  beta + qt(1 - alpha/2, n - p) * sqrt(diag(vcovHC3))
)

colnames(HC3_CI) = colnames(confint(quant_model))
HC3_CI
```

```{r}
library(lmtest)
confint(coeftest(quant_model, vcov. = vcovHC3))
```

```{r}
test = matrix(0, p, 4)
rownames(test) = names(beta)
colnames(test) = c("Estimate", "Std. Error", "t value", "Pr(>|t|)")

test[,1] = beta
test[,2] = sqrt(diag(vcovHC3))
test[,3] = test[,1] / test[,2]
test[,4] = 2 * pt(abs(test[,3]), n - p, lower.tail = FALSE)

test
```

```{r}

hc3_test = coeftest(quant_model, vcov. = vcovHC3)

```

```{r}

results_table <- data.frame(
  Coefficient = rownames(ols_summary$coefficients),
  Estimate = ols_summary$coefficients[, "Estimate"],
  OLS_SE = ols_summary$coefficients[, "Std. Error"],
  HC3_SE = sqrt(diag(vcovHC3)),
  t_HC3 = hc3_test[, "t value"],
  p_HC3 = hc3_test[, "Pr(>|t|)"]
)

results_table
```

The HC3 standard errors are larger for most of the coefficients. 
(Intercept) OLS SE: 1.86166629 HC3 SE: 1.83378787
temperature OLS SE:0.04630927 HC3 SE:0.04588259
pulse OLS SE: 6.97628258 HC3 SE:7.07424389
PCV OLS SE: 0.13902340 HC3 SE: 0.13902340
total_protein OLS SE: 0.03208568 HC3 SE:0.03208568

As the SEs increased, the HC3 t-values got smaller. 
(Intercept): -2.145 to -1.965
temperature: 2.442 to 2.205
pulse: 5.967 to 5.144
PCV: -0.640 to -0.568
total_protein: -1.011 to -0.972.

The p-values got larger. 
(Intercept)	0.0330 to 0.0506 (no longer < 0.05)	
temperature	0.0154 to 0.0284 (still < 0.05)	
pulse	8.82e-09 to 5.67e-07 (still extremely significant)	
PCV	0.5228 to 0.5709 (still NS)	
total_protein	0.3132 to 0.3319 (still NS)

The intercept is the only coefficient whose significance status changed. In the OLS (Intercept) is significant (p = 0.033), but HC3 shows it is only marginally significant (p = 0.0506).

The HC3 CIs are consistently wider:

(Intercept) OLS [-216.74, -9.22]; HC3 [-226.25, 0.29]. Which now covers zero.
temperature: OLS [0.656,6.142]; HC3 [0.362,6.43]
pulse: OLS [0.169,0.336]; HC3 [0.156,0.349]
PCV: OLS [-0.283, 0.144]; HC3 [-0.310, 0.171]
total_protein: OLS [-0.109, 0.0353]; HC3 [-0.112, 0.038]

The qualitative conclusions remained the same for temperature, pulse, PCV and total_protein. 

\qed


\problem
\textit{Proof}.
```{r}
# Weighted LS using temperature
fit = lm(Y ~ temperature, horse)
summary(fit)
X = model.matrix(fit)
n = dim(X)[1]
p = dim(X)[2] - 1
residuals = fit$residuals
plot(rstandard(fit) ~ temperature, horse, ylab = "Standardized
Residuals", pch = 16)
abline(h = 0, lty = 2, lwd = 2)
```
The variance of the residuals appears to be an increasing function of temperature. Since the weights are inversely proportional to the residual variances, we infer that the weights would need to be a decreasing function of temperature.

```{r}
w = rep(1, n)
err = Inf
tol = 1e-10
niter = 0
maxiter = 1e3
while (err > tol & niter < maxiter)
{
  aux = lm(log(residuals^2) ~ log(temperature), horse)
  err = sum(abs(w - exp(-aux$fitted.values))) / sum(w)
  w = exp(-aux$fitted.values)
  WLS = lm(Y ~ temperature, horse, weights = w)
  residuals = WLS$residuals
  niter = niter + 1
}
summary(aux)
summary(WLS)
```
We estimate that $\sigma^2_i\approx1.64\times 10^{-16}\cdot temperature_i^{9.33}$ 
```{r}
plot(WLS$fitted.values, rstandard(WLS), xlab = "Fitted Values", ylab = "Weighted Standardized Residuals", pch = 16)
abline(h = 0, lty = 2, lwd = 2)
```

```{r}
WLS_all = lm(Y ~ temperature + inv_pulse + log_pcv + log_tp, horse, weights = w)
summary(WLS_all)
```
The WLS standard errors are slightly smaller than the corresponding OLS estimate.

```{r}
#t-test and CIs
t_WLS = coef(WLS_all) / summary(WLS_all)$coefficients[, "Std. Error"]
t_WLS

CI_WLS = confint(WLS_all, level = 0.95)
CI_WLS


t_OLS = coef(quant_model) / summary(quant_model)$coefficients[, "Std. Error"]
t_OLS

CI_OLS = confint(quant_model, level = 0.95)
CI_OLS
```
Temperature became significant under WLS, where is was insignificant in OLS. Inv_pulse is highly significant in both. The other coefficients don't change much in significance. The temperature CI under WLS does not include 0, and it does in OLS. WLS gives slightly narrower CIs for all covariates compared to OLS. 

\qed

\problem
\textit{Proof}.
```{r}
# mean comparisons for pain
Q = factor(horse$pain)
n = length(Q)
K = length(levels(Q))
mu = aggregate(Y ~ pain, data = horse, FUN = function(x) mean(x))[,2]
sigma = sqrt(sum((Y - mu[Q])^2) / (n - K))
SE = sigma * sqrt(2 * K / n)
diff = outer(mu, mu, "-")
diff = diff[lower.tri(diff)]
lab = outer(levels(Q), levels(Q), function(x, y) {paste0(x, "-", y)})
names(diff) = lab[lower.tri(lab)]
CI = cbind(diff - qt(1 - alpha / 2, n - K) * SE, diff + qt(1 - alpha / 2, n - K) * SE)
colnames(CI) = c("Lower Bound", "Upper Bound")
print(CI)
```
```{r}
# Tukey’s Honestly Significant Difference Method
# cannot do manual calculations because group sizes are unequal
TukeyHSD(aov(Y ~ pain, horse))
```
Tukey’s Honestly Significant Difference Method (Tukey's HSD) calculates pairwise differences using a critical value from Tukey's studentized range distribution rather than the t-distribution. It can only take positive values and is parametrized by the number of groups (k) and the residual (n-k) degrees of freedom. For each pair of group means $\hat{\mu}_k-\hat{\mu}_\ell$, simultaneous 95% CIs are computed such that FWER $\leq\alpha$ using the upper $\alpha$-quantile of the studentized range divided by $\sqrt2$ and multiplied by the pooled sample standard deviation times the square root of 2 times the number of groups divded by the sample size, eg $\hat{\mu}_k-\hat{\mu}_\ell\pm \frac{R_{K,n-K;\alpha}}{\sqrt2} \cdot S\sqrt{\frac{2K}{n}}$.

The CIs using the t-distribution are generally wider than the narrower Tukey's estimates. Tukey's estimates take into account the studentized range distribution, which adjusts for the fact that multiple comparisons are being made simultaneously while controlling the FWER. Since Tukey's is more conservative, there are less significant comparisons under this method. Only 2 comparisons are significant under Tukey's (continuous-alert and depressed-alert). These CIs do not contain 0. For the CIs using the t-distribution, both of those comparisons are significant, as well as two others (severe-alert and mild-continuous).
\qed

\problem
\textit{Proof}.
```{r}
fit = lm(Y ~ ., data = horse)
summary(fit)

```

```{r}
#Eden Added --- Global Null Testing
Y = log(horse$respiratory_rate)

fit <- lm(
  Y ~ surgery + age + temperature + pulse + inv_pulse +
    temperature_of_extremities + peripheral_pulse + pain +
    peristalsis + abdominal_distension +
    packed_cell_volume + total_protein +
    outcome + surgical_lesion,
  data = horse
)
summary(fit)


alpha = 0.05
pvals <- summary(fit)$coefficients[-1, 4]
m <- length(pvals)

sum(pvals < alpha)
```

```{r}
#Eden Added --- Bonferroni
Bonf = min(pvals)
print(Bonf < alpha / m)
```

```{r}
#Eden Added --- Fishers Combo
Fisher = -2 * sum(log(pvals))
print(Fisher > qchisq(1 - alpha, 2 * m))
```

```{r}
#Eden Added --- Simes
Simes = sum(pvals[order(pvals)] <= (1:m) * alpha / m)
print(Simes > 0)
```

We tested the global null hypothesis that all regression coefficients are equal to zero against the alternative that at least one coefficient is nonzero. The overall F-test from the regression output provides a global hypothesis test that controls the family-wise error rate. The F-statistic of 60.45 with a p-value less than 2.2×10^⁻16 leads us to reject the global null hypothesis. Therefore, there is strong evidence that at least one covariate has a statistically significant effect on the response variable.

Eden added --- After applying both Bonferroni correction to control the family-wise error rate, the Simes Global Test, and the Benjamini–Hochberg procedure to control the false discovery rate at the 5% level, all tested effects remained statistically significant and came to the same conclusion. Bonferroni controls the probability of any false positive and is very conservative, where BH allows some false positives but increases power, and Simes tests the global null that all effects are zero.

\qed

\problem
\textit{Proof}.


```{r}
#Holm-Bonferroni Test
HB = numeric(m)
for (j in 1:m)
{
HB[j] = min(max((m - (1:j) + 1) * pvals[order(pvals)][1:j
]), 1)
}
HB = HB[order(order(pvals))]
HB = p.adjust(pvals, "holm")
sum(HB < alpha)
names(pvals)[HB < alpha]


```

```{r}
#Benjamini-Hochberg Procedure
BH = numeric(m)
for (j in 1:m)
{
BH[j] = min(min(m * pvals[order(pvals)][j:m] / (j:m)), 1)
}
BH = BH[order(order(pvals))]
BH = p.adjust(pvals, "BH")
sum(BH < alpha)
names(pvals)[BH < alpha]

```

```{r}
#Benjamini-Yekutieli Procedure

BY = numeric(m)
for (j in 1:m)
{
BY[j] = min(min(m * sum(1 / (1:m)) * pvals[order(pvals)][
j:m] / (j:m)), 1)
}
BY = BY[order(order(pvals))]
BY = p.adjust(pvals, "BY")
sum(BY < alpha)
names(pvals)[BY < alpha]

```
All three methods yield identical conclusions in this dataset. After adjusting for multiple testing and controlling for other covariates, young horses exhibit a statistically significant difference in respiratory rate compared to the reference age group (old in this case).  This was interesting to see compared to the uncorrected p-values showed 3 coefficients having raw p-values < 0.05 before controlling FWER and FDR showcasing how these controls yield more conservative conclusions. 

\qed